---
title: "Hidden Safety Net of Underutilized Supplemental Insurance in US Agriculture "
date: "`r format(Sys.time(), '%B %d %Y ')`"
output:
  word_document: null
  html_document:
    df_print: paged
---

<!-- preliminaries -->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,fig.height=7, fig.width=8.5, warning = F, message = F)
rm(list=ls(all=TRUE));library('magrittr');library('dplyr');library('tidyverse');library(data.table)
devtools::document()
project_directory <- gsub("data-raw/narratives","",getwd())
mainCrops <- unique(readRDS(paste0(project_directory,"data-raw/draw_list.rds"))[[1]]$commodity_code)
study_environment <- readRDS(paste0(project_directory,"data/study_environment.rds"))
Keep.List<-c("Keep.List",ls())
``` 

**Book of business**
```{r, include = F}
rm(list= ls()[!(ls() %in% c(Keep.List))])
tab01a <- readRDS(paste0(project_directory,"data-raw/output/figure_data/book_of_business.rds"))
tab01b <- readRDS(paste0(project_directory,"data-raw/output/figure_data/agent_summary.rds"))
county_crops <- readRDS(paste0(project_directory,"data-raw/output/figure_data/county_crops.rds"))
unique(tab01a$commodity_code)
net_acre_qty <- round(tab01a[(tab01a$Variable %in% "net_acre_qty" & tab01a$commodity_code %in% "All"),"mean"],2)
liability_amt <- round(tab01a[(tab01a$Variable %in% "liability_amt" & tab01a$commodity_code %in% "All"),"mean"],2)
total_prem <- round(tab01a[(tab01a$Variable %in% "total_prem" & tab01a$commodity_code %in% "All"),"mean"],2)
indem_amt <- round(tab01a[(tab01a$Variable %in% "indem_amt" & tab01a$commodity_code %in% "All"),"mean"],2)
liab_rep <- round(tab01a[(tab01a$Variable %in% "liab_rep" & tab01a$commodity_code %in% "All"),"mean"],2)
lr_cum <- round(tab01a[(tab01a$Variable %in% "lr_cum" & tab01a$commodity_code %in% "All"),"mean"],2)

tab01a$mean <- round(tab01a$mean,2)
tabcrp <- tab01a[tab01a$Variable %in% c("net_acre_qty","liability_amt","lr_cum","liab_rep"),c("commodity_code","Variable","mean")] 
tabcrp <- tidyr::spread(tabcrp,Variable, mean)
tabcrp <- tabcrp[!tabcrp$commodity_code %in% "All",]
tabcrp <- tabcrp[order(-tabcrp$net_acre_qty),]
tabcrp <- tabcrp[1:6,]
tabcrp$text01 <- paste0(tabcrp$net_acre_qty," million acres ($",tabcrp$liability_amt," billion)")
tabcrp$commodity_code[nrow(tabcrp)] <- paste0("and ",tabcrp$commodity_code[nrow(tabcrp)])
tabcrp$text01[nrow(tabcrp)] <- paste0("and ",tabcrp$text01[nrow(tabcrp)])

crop6_list <- tolower(paste0(tabcrp$commodity_code,collapse = ", "))
crop6_liab <- tolower(paste0(tabcrp$text01,collapse = ", "))

tabcrp <- tabcrp[order(tabcrp$lr_cum),]
lr_cum_lo <- tabcrp[1,"lr_cum"]
lr_cum_hi <- tabcrp[nrow(tabcrp),"lr_cum"]
lr_cum_lo_crop <- tolower(tabcrp[1,"commodity_code"])
lr_cum_hi_crop <- tolower(tabcrp[nrow(tabcrp),"commodity_code"])

tabcrp <- tabcrp[order(-tabcrp$liab_rep),]
tabcrp$text02 <- paste0(tabcrp$commodity_code," at ",tabcrp$liab_rep,"%")
tabcrp$text02[nrow(tabcrp)] <- paste0("and ",tabcrp$text02[nrow(tabcrp)])

liab_rep_hi <- tabcrp[1,"lr_cum"]
liab_rep_crop <- tolower(tabcrp[nrow(tabcrp),"commodity_code"])
crop6_liab_rep <- tolower(paste0(tabcrp$text02[2:nrow(tabcrp)],collapse = ", "))

n_crops <- length(unique(county_crops$commodity_name))
n_counties <- nrow(unique(county_crops[c("state_code", "county_code")]))
n_progs <- nrow(unique(county_crops[c("state_code", "county_code","commodity_name")])) 

all_crops <- tolower(unique(county_crops$commodity_name))
all_crops <- all_crops[order(all_crops)]
all_crops[length(all_crops)] <- paste0("and ",all_crops[length(all_crops)])
all_crops <- paste0(all_crops,collapse = ", ")
``` 

Our study focuses on data from 2015-23 for `r n_crops` crops across `r n_counties` counties, encompassing `r n_progs` crop/county programs where SCO and ECO offerings were available.^[The crops included `r all_crops`.] Table S1 presents the descriptive statistics of our dataset. During this period, the annual net insured area averaged `r net_acre_qty` million acres with a total liability of $`r liability_amt` billion, and the program collected premiums totaling $`r total_prem` billion annually. Additionally, it paid out $`r indem_amt` billion in indemnities to producers each year. `r crop6_list` comprised most of this business with insured area (liability) accounting, respectively, for `r crop6_liab`. These crops also led in all other metrics listed in Table S1. The cumulative loss ratio across all crops was `r lr_cum` but varied among the six major crops, ranging from `r lr_cum_lo` for `r lr_cum_lo_crop` to `r lr_cum_hi` for `r lr_cum_hi_crop`. The sample represents `r liab_rep`% of the total non-livestock liability within the FCIP from 2015-23. See Table S2 in the online appendix for the book of business summaries of other crops not shown on Table S1.

**3. Supplemental Coverage Offering and Participation**

```{r, include = F}
rm(list= ls()[!(ls() %in% c(Keep.List))])
data <- as.data.frame(readRDS(paste0(project_directory,"data/supplemental_offering_and_adoption.rds")))


avail_aph_sb <- sum(as.numeric(data$commodity_code %in% 39 & data$commodity_year %in% 2024 & data$avail_aph %in% 1),na.rm=T)
avail_sco_sb <- sum(as.numeric(data$commodity_code %in% 39 & data$commodity_year %in% 2024 & data$avail_aph %in% 1 & data$avail_sco %in% 1),na.rm=T)
avail_eco_sb <- sum(as.numeric(data$commodity_code %in% 39 & data$commodity_year %in% 2024 & data$avail_aph %in% 1 & data$avail_eco90 + data$avail_eco95 >0 ),na.rm=T)

avail_aph <- sum(as.numeric(data$commodity_year %in% 2024 & data$avail_aph %in% 1),na.rm=T)
avail_sco <- sum(as.numeric(data$commodity_year %in% 2024 & data$avail_aph %in% 1 & data$avail_sco %in% 1),na.rm=T)
avail_eco <- sum(as.numeric(data$commodity_year %in% 2024 & data$avail_aph %in% 1 & data$avail_eco90 + data$avail_eco95 >0 ),na.rm=T)

crop_aph <- length(unique(data[(data$commodity_year %in% 2024 & data$avail_aph %in% 1),"commodity_code"]))
crop_sco <- length(unique(data[(data$commodity_year %in% 2024 & data$avail_aph %in% 1 & data$avail_sco %in% 1),"commodity_code"]))
crop_eco <- length(unique(data[(data$commodity_year %in% 2024 & data$avail_aph %in% 1 & data$avail_eco90 + data$avail_eco95 >0),"commodity_code"]))

setDT(data)
crop_avail <- as.data.frame(
  data[commodity_year %in% 2015:study_environment$year_end & commodity_code %in% mainCrops,.(
    avail_aph = sum(avail_aph,na.rm=T),
    avail_sco = sum(avail_sco,na.rm=T),
    avail_eco = sum(as.numeric(avail_eco90 + avail_eco95 > 0),na.rm=T),
    avail_ecosco = sum(as.numeric(avail_sco + avail_eco90 + avail_eco95 > 0),na.rm=T)), 
    by = c("commodity_code")])

crop_avail$avail_sco <- round((crop_avail$avail_sco/crop_avail$avail_aph)*100)
crop_avail$avail_eco <- round((crop_avail$avail_eco/crop_avail$avail_aph)*100)
crop_avail$avail_ecosco <- round((crop_avail$avail_ecosco/crop_avail$avail_aph)*100)

crop_avail <- crop_avail[order(-crop_avail$avail_ecosco),]
``` 


Figure 2 illustrates that SCO and ECO have been available for purchase across nearly all eligible commodity/county programs since their introduction. Eligibility for SCO and ECO is determined by whether the commodity/county program offers an individual-based basic insurance policy – i.e., APH, YP, RP, or RP-HPE. As of `r study_environment$year_end`, out of `r as.character(avail_aph)` eligible county crop programs, `r as.character(avail_sco)` (`r round((avail_sco/avail_aph)*100,2)`%) offered SCO and `r as.character(avail_eco)` (`r round((avail_eco/avail_aph)*100,2)`%) offered ECO. Specifically for the `r crop_aph` commodities eligible  in `r study_environment$year_end`, `r crop_sco` (`r round((crop_sco/crop_aph)*100,2)`%) provided SCO, while only `r crop_eco` (`r round((crop_eco/crop_aph)*100,2)`%) had ECO available. Despite the availability of individual-based basic policies in Hawaii and some areas of Alaska, neither SCO nor ECO are available to producers in these states as of `r study_environment$year_end`. However, both options have been consistently accessible from 2015 to `r study_environment$year_end` in all eligible counties for cotton, `r crop_avail[crop_avail$commodity_code %in% 41,"avail_ecosco"]`% for corn and soybeans, `r crop_avail[crop_avail$commodity_code %in% 11,"avail_ecosco"]`% for wheat, and `r crop_avail[crop_avail$commodity_code %in% 18,"avail_ecosco"]`% for rice (see Figure S1).

```{r, include = F}
rm(list= ls()[!(ls() %in% c(Keep.List))])
tab01a <- as.data.frame(readRDS(paste0(project_directory,"data-raw/output/figure_data/book_of_business.rds")))
fig3 <- as.data.frame(readRDS(paste0(project_directory,"data-raw/output/figure_data/supplemental_coverage_temp.rds")))
fig3$plan <- ifelse(grepl("SCO",as.character(fig3$plan)),"SCO","ECO")

net_acre_qty <- round(tab01a[(tab01a$Variable %in% "net_acre_qty" & tab01a$commodity_code %in% "All"),"mean"],2)
edr_acre_qty <- round(tab01a[(tab01a$Variable %in% "edr_acre_qty" & tab01a$commodity_code %in% "All"),"mean"],2)
county_crops <- readRDS(paste0(project_directory,"data-raw/output/figure_data/county_crops.rds"))

setDT(fig3)
fig3 <- rbind(fig3[plan %in% "SCO" & commodity_year %in% c(2015,study_environment$year_end)],
                 fig3[plan %in% "ECO" & commodity_year %in% c(2021,study_environment$year_end)]) %>% 
  group_by(cropname) %>% mutate(commodity_year_beg = min(commodity_year,na.rm=T)) %>% as.data.frame(.)
setDT(fig3)
fig3 <- as.data.frame(
  fig3[type %in% "Adoption as a percentage of eligible insured acres",.(
    value = round(mean(value,na.rm=T),2)), 
    by = c("cropname","commodity_code","plan","commodity_year")])
fig3$commodity_year <- ifelse(fig3$commodity_year %in% study_environment$year_end,"v1","v0")
fig3 <- tidyr::spread(fig3,commodity_year, value)
fig3$x <- round(fig3$v1-fig3$v0,2)
fig3

sixcrp <- fig3[fig3$cropname %in% c("Rice","Wheat","Corn","Soybeans","Cotton","Sorghum"),c("cropname","plan","v1")]
sixcrp <- tidyr::spread(sixcrp,plan, v1)
sixcrp$name <- tolower(sixcrp$cropname)
sixcrp$rate <- paste0(sixcrp$SCO,"% [",sixcrp$ECO,"%]")
sixcrp$rate[nrow(sixcrp)] <- paste0("and ",sixcrp$rate[nrow(sixcrp)])
sixcrp$name[nrow(sixcrp)] <- paste0("and ",sixcrp$name[nrow(sixcrp)])

sixgwt <- fig3[fig3$cropname %in% c("Rice","Wheat","Corn","Soybeans","Cotton","Sorghum"),c("cropname","plan","x")]
sixgwt <- tidyr::spread(sixgwt,plan, x)
sixgwt$cropname <- tolower(sixgwt$cropname)
sixgwt <- sixgwt[order(-sixgwt$SCO),]
sixgwt$rate <- sixgwt$SCO
sixgwt$cropname[nrow(sixgwt)-1] <- paste0("and ",sixgwt$cropname[nrow(sixgwt)-1])
sixgwt$rate[nrow(sixgwt)-1] <- paste0("and ",sixgwt$rate[nrow(sixgwt)-1])
sixgwt

crop_avail <- unique(as.data.frame(readRDS(paste0(project_directory,"data-raw/output/figure_data/crop_availability.rds"))))
crop_avail <- crop_avail[!crop_avail$commodity_name %in% c("Rice","Wheat","Corn","Soybeans","Cotton","Sorghum","All Commodities"),]
crop_avail <- crop_avail[order(!crop_avail$sco_use),]
top5crp <- crop_avail[crop_avail$sco_use >= 5,]
top5crp$txt <- paste0(tolower(top5crp$commodity_name)," (",round(top5crp$sco_use,2),")")
top5crp$txt[nrow(top5crp)] <- paste0("and ",top5crp$txt[nrow(top5crp)])

lowcrp <- crop_avail[crop_avail$sco_use < 1,]
lowcrp$txt <- tolower(lowcrp$commodity_name)
lowcrp$txt[nrow(lowcrp)] <- paste0("and ",lowcrp$txt[nrow(lowcrp)])

``` 

Despite their widespread availability, usage of SCO and ECO is relatively small compared to the number of eligible county-crop program offers, Particularly, Table S1 shows that of the `r net_acre_qty` million acres with SCO/ECO availability that were annually enrolled in the FCIP from 2015-`r study_environment$year_end-2000`, only `r edr_acre_qty` million acres (i.e., `r round((edr_acre_qty/net_acre_qty)*100,2)`% of the total) were enrolled in a supplemental program. The portion of insured eligible acres annually enrolled in SCO [ECO] in `r study_environment$year_end` for `r paste0(sixcrp$name,collapse = ", ")`, were `r paste0(sixcrp$rate,collapse = ", ")`, respectively. Figure S1 shows that while SCO adoption declined by `r abs(as.numeric(sixgwt$rate[6]))` percentage points among `r sixgwt$cropname[6]` producers from 2015-`r study_environment$year_end-2000`, `r paste0(sixgwt$cropname[1:(nrow(sixgwt)-1)],collapse = ", ")` experienced growth estimated at `r paste0(sixgwt$rate[1:(nrow(sixgwt)-1)],collapse = ", ")` percentage points, respectively. Aside from these major crops, notable exceptions of SCO adoption that is more than 5% of eligible acres from 2015-`r study_environment$year_end-2000` include `r paste0(top5crp$txt,collapse = ", ")`. On the contrary, `r paste0(lowcrp$txt,collapse = ", ")` recorded SCO adoption rates of less than 1% of eligible acres from 2015-`r study_environment$year_end-2000` (see Table S3 in the online appendix).

**Agent characteristics**
```{r, include = F}
rm(list= ls()[!(ls() %in% c(Keep.List))])
tab01a <- as.data.frame(readRDS(paste0(project_directory,"data-raw/output/figure_data/book_of_business.rds")))
tab01b <- as.data.frame(readRDS(paste0(project_directory,"data-raw/output/figure_data/agent_summary.rds")))
county_crops <- as.data.frame(readRDS(paste0(project_directory,"data-raw/output/figure_data/county_crops.rds")))

obs_n <- as.character(round(tab01b[(tab01b$Variable %in% "obs_n" & tab01b$commodity_code %in% "All"),"mean"],0))
Corn_n <- as.character(round(tab01b[(tab01b$Variable %in% "obs_n" & tab01b$commodity_code %in% "Corn"),"mean"],0))
Soybeans_n <- as.character(round(tab01b[(tab01b$Variable %in% "obs_n" & tab01b$commodity_code %in% "Soybeans"),"mean"],0))
Wheat_n <- as.character(round(tab01b[(tab01b$Variable %in% "obs_n" & tab01b$commodity_code %in% "Wheat"),"mean"],0))
Cotton_n <- as.character(round(tab01b[(tab01b$Variable %in% "obs_n" & tab01b$commodity_code %in% "Cotton"),"mean"],0))
Rice_n <- as.character(round(tab01b[(tab01b$Variable %in% "obs_n" & tab01b$commodity_code %in% "Rice"),"mean"],0))
Sorghum_n <- as.character(round(tab01b[(tab01b$Variable %in% "obs_n" & tab01b$commodity_code %in% "Sorghum"),"mean"],0))

cov_lvl <- round(tab01b[(tab01b$Variable %in% "coverage_level_percent" & tab01b$commodity_code %in% "All"),"mean"]*100,0)
comm_amt <- round(tab01b[(tab01b$Variable %in% "insured_acres" & tab01b$commodity_code %in% "All"),"mean"]*1000,0)
premium_rate <- round(tab01b[(tab01b$Variable %in% "premium_rate" & tab01b$commodity_code %in% "All"),"mean"]*100,0)
Subsidy_Percent <- round(tab01b[(tab01b$Variable %in% "Subsidy_Percent" & tab01b$commodity_code %in% "All"),"mean"]*100,0)
rep_samp <- round((tab01b[(tab01b$Variable %in% "agent_liab" & tab01b$commodity_code %in% "All"),"mean"]/
                            tab01b[(tab01b$Variable %in% "sob_liab" & tab01b$commodity_code %in% "All"),"mean"])*100,2)

n_progs <- nrow(unique(county_crops[c("state_code", "county_code","commodity_name")])) 

tabcrp <- tab01b[tab01b$Variable %in% c("agent_liab","sob_liab"),c("commodity_code","Variable","mean")] 
tabcrp <- tidyr::spread(tabcrp,Variable, mean)
tabcrp <- tabcrp[tabcrp$commodity_code %in% c("Rice","Wheat","Corn","Soybeans","Cotton","Grain Sorghum"),]
tabcrp$rep_samp <- round((tabcrp$agent_liab/tabcrp$sob_liab)*100,2)
tabcrp <- tabcrp[order(-tabcrp$rep_samp),]
tabcrp$commodity_code <- tolower(gsub("Grain ","",tabcrp$commodity_code))
tabcrp$text01 <- paste0(tabcrp$commodity_code," at ",tabcrp$rep_samp,"%")
tabcrp$text01[nrow(tabcrp)] <- paste0("and ",tabcrp$text01[nrow(tabcrp)])

``` 

The lower part of Table S1 presents selected descriptive statistics for our sample agents, totaling `r obs_n` across all crops. Among the six major crops, the number of agents is as follows: corn with `r Corn_n` agents, soybeans with `r Soybeans_n`, wheat with `r Wheat_n`, cotton with `r Cotton_n`, sorghum with `r Sorghum_n`, and rice with `r Rice_n`. The most utilized basic insurance plan among these agents is Revenue Protection (RP), featuring a weighted average coverage level of `r cov_lvl`% and an average insured area of `r comm_amt` acres per agent. On average, these agents pay a premium of `r premium_rate` cents per dollar of liability and receive a subsidy of `r Subsidy_Percent` cents per dollar of premium for the basic policy. The agents included in our simulations represent `r rep_samp`% of the total liability of the `r n_progs` crop/county programs where SCO and ECO offerings were available from 2015-`r study_environment$year_end-2000`. Among the six major crops, the simulation representation was highest for `r tabcrp$commodity_code[1]` at `r tabcrp$rep_samp[1]`%, followed by `r paste0(tabcrp$text01[2:nrow(tabcrp)],collapse = ", ")`. 










